---
title: "MMP"
author: "Neal Marquez"
date: "May 20, 2018"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
rm(list=ls())

library(readr)
library(dplyr)
library(ggplot2)

# Note that in the data 9999 often refers to unknown while
# 8888 refers to somme measure being not applicable
mmpDF <- read_csv("./data/pers161.csv", na = c("", "NA", 9999))
mmpDF %>%
    select(surveyyr, commun, surveypl) %>%
    unique %>%
    group_by(commun) %>%
    summarize(Count=n()) %>%
    select(Count) %>%
    unlist %>%
    table
```

```{r}
comms <- c(9, 11, 37, 44, 45)

mmpDF %>%
    select(commun, weight, surveyyr, surveypl) %>%
    group_by(commun, surveyyr, surveypl) %>%
    summarize(weight=mean(weight), n=n()) %>%
    mutate(hhs=weight*n)
```

```{r}
mmpDF %>%
    filter(surveypl==1) %>%
    filter(age >=15 & commun %in% comms & sex == 1) %>%
    mutate(everMig=usyr1 != 8888) %>%
    mutate(ustrips=ifelse(ustrips==8888, NA, ustrips)) %>%
    group_by(commun) %>%
    ungroup %>%
    select(everMig, ustrips, commun, weight) %>%
    group_by(commun) %>%
    summarize(
        N=n(), 
        Weight=mean(weight),
        eMig=100 * mean(everMig, na.rm=T), 
        Trips=mean(ustrips, na.rm=T),
        missT=sum(is.na(ustrips)),
        missM=sum(is.na(everMig)))

mmpDF %>%
    filter(age >=15 & commun %in% comms & sex == 1) %>%
    mutate(ustrips=ifelse(ustrips==8888, NA, ustrips)) %>%
    group_by(commun) %>%
    filter(surveyyr == min(surveyyr)) %>%
    ungroup %>%
    select(ustrips, commun, weight) %>%
    mutate(weight=weight/sum(weight*!is.na(ustrips), na.rm=T)) %>%
    summarize(
        wTrips=round(sum(ustrips*weight, na.rm=T), 2),
        Trips=round(mean(ustrips, na.rm=T), 2))

mmpDF %>%
    filter(age >=15 & commun %in% comms & sex == 1) %>%
    mutate(everMig=ifelse(is.na(usyr1), NA, usyr1 != 8888)) %>%
    group_by(commun) %>%
    filter(surveyyr == min(surveyyr)) %>%
    ungroup %>%
    select(everMig, commun, weight) %>%
    mutate(weight=weight/sum(weight*!is.na(everMig), na.rm=T)) %>%
    summarize(
        wTrips=round(100*sum(everMig*weight, na.rm=T), 1),
        Trips=round(100*mean(everMig, na.rm=T), 1))
```

## Plot first age at migration for males over 15 

Here we try to replicate the plots from the paper "U.S. Immigration Reform and the Migration Dynamics of Mexican Males" by Khulan Altangerel and Jan C. van Ours. As part of their analysis they examine males risk of migration and look at how the probability of having never migrated decreases over age. Their analysis only goes up to the 154 study so we will exclude all results before community 154's first data collection in the year 2015.

```{r}
DFmig <- mmpDF %>%
    # not us born
    filter(usborn == 2) %>%
    # remove all observations before commun 154
    filter(commun <=154 & surveyyr <= 2015) %>%
    # remove unknown data
    filter(!is.na(usyr1)) %>%
    # recode never migrated usyr1 to be the new NA
    mutate(usyr1=ifelse(usyr1 == 8888, NA, usyr1)) %>%
    # create age at time of interview
    mutate(age=surveyyr-yrborn) %>%
    # create age at first migration
    mutate(usag1=usyr1-yrborn) %>%
    # we only want male respondents and only if they are older than 15
    filter(sex==1 & age >=15) %>%
    # we only count migrations if they occured after the age of 14(>= 15)
    filter(is.na(usag1) | usag1 >= 15) %>%
    # select only a subset of the categories
    select(usag1, weight, age) %>%
    # estimate the age to which an individual contributes to teh denominator
    rowwise %>%
    mutate(ycontrib=min(age, usag1, na.rm=T)) %>%
    ungroup

data.frame(
    pmig=sapply(14:35, function(x){
        DFmig %>%
            # select only the individuals who contributed up to that year
            filter(ycontrib >= x) %>%
            # calculate the weights for denominator so they sum to one
            mutate(weight=weight/sum(weight)) %>%
            #mutate(weight=1/nrow(.)) %>%
            # lets get the contributors to numerator
            filter(usag1 == x) %>%
            select(weight) %>%
            unlist %>%
            sum
        })) %>%
    mutate(age=14:35) %>%
    mutate(pnevermig=cumprod(1-pmig)) %>%
    ggplot(aes(x=age, y=pnevermig)) +
    geom_line() + 
    theme_bw() + 
    ylim(c(0.5, 1)) +
    labs(x="Age", y="Probability of Never Migrated") +
    ggtitle("Weighted Survival Rates of First Migration")
```

```{r}
commDF <- read_csv("./data/commun161.csv")
sort(names(commDF))
```